---
title: "Oral Discussion Example Program"
author: "Justin Post"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(haven)
```

# Introduction

The purpose of this document is to read some data into R and explore it briefly. Below you'll find R **code chunks** with code followed by output from that code. We use the `readr`, `readxl`, and `haven` packages as needed.

# Read in Data

Let's begin by reading in data from a URL. 

```{r}
air <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/AirQuality.csv")
```

Now we'll read in data from a local file.

```{r}
#This data is available on the Moodle site (Resources & Information section)
chickens <- read_excel("Chickens.xlsx") 
```

Lastly, let's bring in data from SAS.

```{r}
smoke <- read_sas("https://www4.stat.ncsu.edu/~online/datasets/smoke2003.sas7bdat")
```

# Manipulate Data

Often we'll need to manipulate the data. We can use some base R commands and functions such as `[` to do so.

## Air Quality

Let's manipulate the air quality data.

```{r}
air[1:10, c("RH", "Date")]
x <- c(1, 5, 10)
air[x, ]
```


## Chickens

Let's manipulate the chickens data.

```{r}
chickens[, 1:2]
chickens$weightGain
```

## Chickens

Let's manipulate the smoke data.

```{r}
smoke$PacksPerDay
y <- 1:400
smoke[-y, 1:4]
```


# Together

Put create an R object with info about this data.

```{r}
data_names <- c("air", "chickens", "smoke")
data_files <- list(air, chickens, smoke)
names(data_files) <- data_names

data_files[["air"]]
data_files[[3]]

columns <- list(colnames(air), colnames(chickens), colnames(smoke))
columns[3]
columns[[3]]
```

